---
title: "Practical Machine Learning Final Project"
output: html_document
---
How I built the model:
First, the question we want to answer in this project with is what data collected by different fitness devices can help us to predict the manner in which our participants exercise? 
The data used here are sourced from http://groupware.les.inf.puc-rio.br/har. 
```{r}
###################
#   Read in data
###################
library(caret)
library(ggplot2)
library(knitr)
training=read.csv(file="pml-training.csv",header=TRUE, sep=",")
testing=read.csv(file="pml-testing.csv",header=TRUE, sep=",")
```
First, take a look at the columns: 
```{r}
str(training)  #Check for the column names and their value types
```
The first six columns do not seem to be relevant in our prediction, so I decided to delete them in the model.
```{r}
deleteCol= 1:6   #I decided to delete the first 6 columns as they don't appear to be relevant
```
Then, I check for columns with a lot of NA values, and add those column numbers to the to-be-deleted list. Then delete all the above. 
```{r}
NAcolInd=(union(which(colSums(is.na(training))>0),which(colSums(is.na(testing))>0)))
deleteCol=c(deleteCol,NAcolInd[order(NAcolInd)])
training=training[,-deleteCol]  #Delete these columns
```
Next, set training and testing data, with 70% of the data in training.

```{r}
inTrain<-createDataPartition(y=training$classe,p=0.7,list=FALSE)
train<-training[inTrain,]
tests<-training[-inTrain,]
```
Then I wanted to check on variables that have high correlations (>0.9): 
```{r}
nrows=dim(training)[1]
ncols=dim(training)[2]
corre=cor(training[2:nrows,1:(ncols-1)])
diag(corre)=0
which(corre > 0.9, arr.ind=TRUE,useNames=TRUE)
```
It appears that "accel_belt_y", "roll_belt" have high correlations with each other and the total variable "total_accel_belt", so I wanted to check out their relationships using a scatterplot matrix:
```{r}
library(car)
scatterplot.matrix(~accel_belt_y+roll_belt+total_accel_belt, data=training,
  	main="Scatterplot matrix of three belt variables")
```


There appears to be some relationships between the variables, but I decided to let the algorithm to select which of these to use in the end model. 

The algorithm I decided to use is random forest, because of its comparably higher accuracy. I ran the regression on all the remaining variables (after eliminating all NA columns and irrelevant columns). 

For cross validation, I used K-fold cross validation, with K=10, and repeat it 3 times. This is because K-fold cross validation seems to work well, and K between 5 and 10 has been proved to be effective. 
```{r}
control<-trainControl(method="repeatedcv",number=10, repeats=3)
```
Train the model: 
```{r}
set.seed(32323)
modelrf<-train(classe ~ .,data=train, preProcess=c("center","scale"),method="rf",trControl=control)
test.pred<-predict(modelrf,tests)
```
Then, get a report of estimated out of sample errors， including accuracy and OOB estimated error. 
```{r}
modelrf$finalModel
confusionMatrix(tests$classe,test.pred)
```

Finally, predict the 20 cases:
```{r}
answer.pred1<-predict(modelrf,testing)
Index<-1:20
results1=data.frame(Index,answer.pred1)
kable(results1)
```
In     conclusion, I built the model using random forest, with preProcessed data. I used K-fold cross validation with K=10. Then I estimated out of sample error based on that. 